---
title: "Adverse event"
output: html_document
date: "`r Sys.Date()`"
---
```{r,include=FALSE}
library(tidyverse)
library(tableone)
library(writexl)
library(gtsummary)
library(survival)
library(ggplot2)
library(readxl)
library(ggpubr)
library(splines)
library(xlsx)
library(lattice)
library(mice)
library(DescTools)
library(survminer)
library(finalfit)
library(tidycmprsk)
library(ggsurvfit)
library(car)
library(cobalt)
library(pammtools)
library(adjustedCurves)
library(rms)
library(riskRegression)
library(prodlim) 
library(dplyr)
library(bodycompref)
library(musclerefdata)
library(transplantr)
library(sqldf)
library(table1)

final_master <- read_xlsx("/Users/to909/Desktop/Terlipressin/final_master_01282026.xlsx")

final_master <- final_master %>%
  mutate(across(everything(), ~{
    x_char <- as.character(.x)
    if (all(grepl("^[-+]?[0-9]*\\.?[0-9]+$", x_char[!is.na(x_char)]))) {
      as.numeric(x_char)
    } else {
      .x  # leave unchanged
    }
  }))

final_master <- final_master %>%
  mutate(
    BMI = aki_weight / (aki_height/100)^2,
    rrt_90days=ifelse(is.na(days_to_rrt),0,ifelse(days_to_rrt<=90,1,0)),
    lt_90days=ifelse(is.na(time_to_transplant),0,ifelse(time_to_transplant<=90,1,0)),
   rrt_status=case_when((days_to_rrt -  time_to_transplant) == 0 ~ 0,
                         (days_to_rrt -  time_to_transplant) < 0 ~ 1,
                         (days_to_rrt -  time_to_transplant) > 0 ~ 2),
   abdominal_ae=if_else(terli_ae_abdpain ==1 |terli_ae_diarrhea == 1 | terli_ae_nausea==1,1,0 ),
   serious_ae=if_else(terli_ae_respfail==1 | terli_ae_brady==1 | terli_7==1 |terli_8==1,1,0),
    meld_fda_value = coalesce(
      day0_meld,
      admit_meld,
      admit_meld_3
    ),
   meld_fda=if_else(meld_fda_value>35,1,0)
  )
   
  final_master$time_to_death_status_90days_cmp <- as.factor(final_master$time_to_death_status_90days_cmp)
```


# Table 1 function
```{r,echo=FALSE,warning=FALSE,message=FALSE}

create_table_one <- function(all_vars,
                             data,
                             strata,
                             output,
                             cat_vars = NULL,
                             exact = NULL) {  # NEW optional input
  
  # 1) Determine Numerical Vars based on presence of cat_vars
  if (is.null(cat_vars)) {
    numerical_vars <- all_vars
  } else {
    numerical_vars <- setdiff(all_vars, cat_vars)
  }
  
  # 2) Generate the TableOne object
  T1_after <- tableone::CreateTableOne(
    vars       = all_vars,
    strata     = strata,
    data       = data,
    factorVars = cat_vars,   # accepts NULL safely
    addOverall = TRUE,
    includeNA  = FALSE
  )
  
  # 3) Print TableOne object and convert to data frame
  T1_after_print <- print(
    T1_after,
    showAllLevels = TRUE,
    exact         = exact,          # accepts NULL safely
    nonnormal     = numerical_vars,
    missing       = TRUE,
    quote         = FALSE,
    noSpaces      = TRUE,
    printToggle   = FALSE
  )
  
  # 4) Write the table as CSV
  write.csv(T1_after_print, file = output, row.names = TRUE)
}
```


# table1_criteria_2.csv
```{r,echo=FALSE}

all_vars <-  c(
  "any_ae",
  "terli_ae_abdpain",
  "terli_ae_nausea",
  "terli_ae_diarrhea",
  "terli_ae_volover",
  "terli_ae_brady",
  "terli_7",
  "terli_8",
  "terli_ae_respfail"
)

categorical_vars <-  c(
  "any_ae",
  "terli_ae_abdpain",
  "terli_ae_nausea",
  "terli_ae_diarrhea",
  "terli_ae_volover",
  "terli_ae_brady",
  "terli_7",
  "terli_8",
  "terli_ae_respfail"
)


exact_vars <- c(

  "terli_ae_abdpain",
  "terli_ae_nausea",
  "terli_ae_diarrhea",
  "terli_ae_volover",
  "terli_ae_brady",
  "terli_7",
  "terli_8"
)

create_table_one(
  all_vars = all_vars,
  cat_vars = categorical_vars,
  data     = final_master,
  exact= exact_vars,
  strata   = "criteria_2",
  output   = '/Users/to909/Desktop/Terlipressin projects/Adverse event/table1_criteria_2.csv'
)

```




# table table1_criteria_3.csv
```{r,echo=FALSE}

all_vars <-  c(
  "any_ae",
  "terli_ae_abdpain",
  "terli_ae_nausea",
  "terli_ae_diarrhea",
  "terli_ae_volover",
  "terli_ae_brady",
  "terli_7",
  "terli_8",
  "terli_ae_respfail"
)

categorical_vars <-  c(
  "any_ae",
  "terli_ae_abdpain",
  "terli_ae_nausea",
  "terli_ae_diarrhea",
  "terli_ae_volover",
  "terli_ae_brady",
  "terli_7",
  "terli_8",
  "terli_ae_respfail"
)


exact_vars <- c(

  "terli_ae_abdpain",
  "terli_ae_nausea",
  "terli_ae_diarrhea",
  "terli_ae_volover",
  "terli_ae_brady",
  "terli_7",
  "terli_8"
)

create_table_one(
  all_vars = all_vars,
  cat_vars = categorical_vars,
  data     = final_master,
  exact= exact_vars,
  strata   = "criteria_3",
  output   = '/Users/to909/Desktop/Terlipressin projects/Adverse event/table1_criteria_3.csv'
)

```

```{r,echo=FALSE}
create_table_one(
  all_vars = all_vars,
  cat_vars = categorical_vars,
  data     = final_master,
  exact= exact_vars,
  strata   = "criteria_4",
  output   = '/Users/to909/Desktop/Terlipressin projects/Adverse event/table1_criteria_4.csv'
)

```

```{r,include=FALSE}
categorical_vars <- c(
  "admit_admitservice",
  "admit_admitroute",
  "sex_male",
  "race",
  "hispanic",
  "insurance_type",
  "cirrhosis_etiology",
  "etoh_hepatitis",
  "hegrade_admit",
  "admit_ascites",
  "htn_medical_hx",
  "dm_medical_hx",
  "ckd_medical_hx",
  "cad_medical_hx",
  "nsbb_preadmission",
  "nsbb_type_preadmission",
  "ace_arb_preadmission",
  "nsaid_preadmission",
  "diuretics_preadmission",
  "sbp_ppx_preadmission",
  "lactulose_preadmission",
  "rifaximin_preadmission",
  "mido_preadmission",
  "ascites_previous_comp",
  "refascites_previous_comp",
  "he_previous_comp",
  "sbp_previous_comp",
  "nonsbp_previous_comp",
  "phtn_bleed_previous_comp",
  "hcc_previous_comp",
  "tips_previous_comp",
  "baseline_scr_outpt",
  "aki_previous_90days",
  "transfer_aki",
  "aki_precipitant_nsbpinf",
  "he_aki",
  "rrt_aki",
  "mv_aki",
  "pressors_aki",
  "chest_imaging",
  "chest_imaging_pulmedema",
  "chest_imaging_effusion",
  "ecg",
  "echo",
  "albumin_before_terli",
  "crystalloid_before_terli",
  "nsbb_before_terli",
  "mido_before_terli",
  "oct_before_terli",
  "ne_before_terli",
  "gihep_consult",
  "neph_consult",
  "pall_consult",
  "code_status",
  "ua_hrs",
  "ua_na_hrs",
  "ua_cr_hrs",
  "ua_urea_hrs",
  "ua_granularcasts_hrs",
  "ua_epi_cells_hrs",
  "location_terli_day0",
  "terli_route_day0",
  "tele_terli_day0",
  "he_terli_day0",
  "infection_terli_day0",
  "typeinfection_terli_day0",
  "hrs_responders",
  "dosechange_terli_day1",
  "dosechange_terli_day2",
  "dosechange_terli_day3",
  "dosechange_terli_day4",
  "dosechange_terli_day5",
  "dosechange_terli_day6",
  "dosechange_terli_day7",
  "dosechange_terli_day8",
  "dosechange_terli_day9",
  "dosechange_terli_day10",
  "dosechange_terli_day11",
  "dosechange_terli_day12",
  "discont_aki_stage",
  "aclf_grade",
  "HRS_Criteria",
  "criteria_1",
  "criteria_2",
  "criteria_3",
  "criteria_4",
  "akin",
  "transplant_slkt",
  "transplant_lt",
  "rrt",
  "rrt_2",
  "rrt_modality",
  "rrt_continuous",
  "hd_discharge",
  "disposition",
  "liver_f",
  "brain_f",
  "coagulation_f",
  "circulatory_f",
  "respiratory_f",
  "kidney_f",
  "baseline_scr_120to365",
  "aki_events_previous90days",
  "prbc_trans_before_terli",
  "time_to_death_status",
   "terli_ae_abdpain", 
    "terli_ae_nausea" ,
    "terli_ae_diarrhea" ,
    "terli_ae_volover" ,
    "terli_ae_brady" ,
    "terli_7",
    "terli_8",
  "terli_ae_respfail","any_ae","listed_transplant","dosechange_status","hrs_rebound","race_2","HRS_reversal_new",
  "hrs_responders_cat_2","time_to_death_status_90days","time_to_death_status_90days_cmp","ckd_baseline","HRS_criteria_3"
)


# Vector of numerical variables
numerical_vars <- c(
  "age",
  "scr_admit",
  "tbili_admit",
  "inr_admit",
  "sodium_admit",
  "albumin_admit",
  "lactate_admit",
  "nsbb_dose_preadmission",
  "mido_dose_preadmission",
  "baseline_scr",
  "baseline_scr_days_to_admit",
 

  "transfer_aki_days_toadmit",
  "days_admit_to_aki",
  "aki_weight",
  "aki_height",
  "aki_hr",
  "aki_sbp",
  "aki_dbp",
  "aki_map",
  "aki_pao2",
  "fio2_aki",
  "scr_aki",
  "bun_aki",
  "albumin_aki",
  "inr_aki",
  "wbc_aki",
  "plt_aki",
  "ecg_qtc",
  "days_to_echo",
  "echo_ef",
  "echo_lateral_e",
  "echo_trpeak",
  "echo_gls",
  "echo_earatio",
  "echo_rvsp",
  "echo_pasp",
  "echo_lavi",
  "echo_tr",
  "totalalbumin_before_terli",
  "albvolume_before_terli",
  "saline_vol_before_terli",
  "lr_volume_before_terli",
  "other_cryst_before_terli",

  "prbc_qaunt_before_terli",
  "totalvolume_before_terli",
  "total_uo_before_terli",
  "days_first_mido_before_terli",
  "days_total_mido_before_terli",
  "mido_maxdose_before_terli",
  "days_first_oct_before_terli",
  "days_total_oct_before_terli",
  "oct_maxdose_before_terli",
  "days_first_ne_before_terli",
  "days_total_ne_before_terli",
  "days_to_gihep_consult",
  "days_to_neph_consult",
  "days_to_pall_consult",
  "days_ua_hrs",
  "days_ua_na_hrs",
  "ua_na_value_hrs",
  "days_ua_cr_hrs",
  "ua_cr_value_hrs",
  "days_ua_urea_hrs",
  "ua_urea_value_hrs",
  "feurea_hrs",
  "ua_rbc_hrs",
  "uaprotein_hrs",
  "temp_terli_day0",
  "hr_terli_day0",
  "pao2_terli_day0",
  "fio2_terli_day0",
  "sbp_terli_day0_time0",
  "map_terli_day0_time0",
  "chloride_terli_day0",
  "bicarb_terli_day0",
  "scr_terli_day0",
  "bun_terli_day0",
  "ast_terli_day0",
  "alt_terli_day0",
  "bili_terli_day0",
  "alb_terli_day0",
  "inr_terli_day0",
  "wbc_terli_day0",
  "hgb_terli_day0",
  "Discont_Day",
  "na_terli_day0",
  "days_to_terli_initiation_day0",
  "clif_c_score",
  "admit_meld",
  "admit_meld_na",
  "admit_meld_3",
  "admit_meld_3_raw",
  "day0_meld",
  "day0_meld_na",
  "day0_meld_3",
  
  "time_to_transplant",
  "time_to_death_status",
  "time_to_death_90days",
  "cystatinc_admit","cystatinc_aki","cystatinc_terli_day0",
  "urine_pc","fena","fena_hrs","time_to_death_90days","total_albumin","mido_maxdose_before_terli "
)




categorical_vars <- intersect(categorical_vars, names(final_master))
numerical_vars <- intersect(numerical_vars, names(final_master))
all_vars <- c(numerical_vars,categorical_vars)

```

# table1_any_ae.csv
```{r,echo=FALSE}



create_table_one(
  all_vars = all_vars,
  cat_vars = categorical_vars,
  data     = final_master,
  strata   = "any_ae",
  output   = '/Users/to909/Desktop/Terlipressin projects/Adverse event/table1_any_ae.csv'
)


```


# table1_abdominal_ae.csv
```{r,echo=FALSE}



create_table_one(
  all_vars = all_vars,
  cat_vars = categorical_vars,
  data     = final_master,
  strata   = "abdominal_ae",
  output   = '/Users/to909/Desktop/Terlipressin projects/Adverse event/table1_abdominal_ae.csv'
)

```

# table1_serious_ae.csv
```{r,echo=FALSE}



create_table_one(
  all_vars = all_vars,
  cat_vars = categorical_vars,
  data     = final_master,
  strata   = "serious_ae",
  output   = '/Users/to909/Desktop/Terlipressin projects/Adverse event/table1_serious_ae.csv'
)

```

# table1_terli_ae_respfail.csv
```{r,echo=FALSE}

create_table_one(
  all_vars = all_vars,
  cat_vars = categorical_vars,
  data     = final_master,
  strata   = "terli_ae_respfail",
  output   = '/Users/to909/Desktop/Terlipressin projects/Adverse event/table1_terli_ae_respfail.csv'
)

```

# column: any_ae 
```{r,echo=FALSE,warning=FALSE,message=F}

table1(~aclf_grade|any_ae,final_master)


```

# logistic regression outcome: any_ae 
```{r,echo=FALSE,warning=FALSE,message=F}

final_master <- final_master %>% mutate_at(c("aclf_grade", "respiratory_f"),as.factor)

final_master  %>%  select( any_ae, scr_terli_day0, respiratory_f
    
) %>%
  tbl_uvregression(
    method = glm,
    method.args = list(family = binomial),
    y = any_ae,
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  ) %>%
  modify_caption("**Table 1. Univariable logistic regression for any adverse event**") %>%
  bold_p(0.05) %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() %>% add_n()

```




#Unviariate cmp:  time_to_death_status_90days_cmp failcode=1 death:  
```{r,echo=FALSE,warning=FALSE,message=F}


final_master %>% 
  select(any_ae, terli_ae_abdpain, terli_ae_nausea, terli_ae_diarrhea, terli_ae_volover, terli_ae_brady, terli_7, terli_8, terli_ae_respfail, serious_ae,time_to_death_90days_cmp,time_to_death_status_90days_cmp) %>%
  tbl_uvregression(
    method = tidycmprsk::crr,
    method.args = list(failcode=1,cencode=0),
    y = Surv(time_to_death_90days_cmp,time_to_death_status_90days_cmp),
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  )

```

#1. Multiviariate cmp:  time_to_death_status_90days_cmp failcode=1 death:any_ae+ age+race+hispanic+ day0_meld
```{r,echo=FALSE,warning=FALSE,message=F}
tidycmprsk::crr(Surv(time_to_death_90days_cmp,time_to_death_status_90days_cmp) ~any_ae+ age+race+hispanic+ day0_meld,failcode=1,cencode=0, data = final_master) %>% tbl_regression(exp = TRUE) %>% add_n()%>%
  bold_p(0.05) %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() 

```
#2. Multiviariate cmp:  time_to_death_status_90days_cmp failcode=1 death:any_ae+ age+race+hispanic
```{r,echo=FALSE,warning=FALSE,message=F}
tidycmprsk::crr(Surv(time_to_death_90days_cmp,time_to_death_status_90days_cmp) ~any_ae+ age+race+hispanic,failcode=1,cencode=0, data = final_master) %>% tbl_regression(exp = TRUE) %>% add_n()%>%
  bold_p(0.05) %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() 

```
#3. Multiviariate cmp:  time_to_death_status_90days_cmp failcode=1 death :serious_ae+ age+race+hispanic+ admit_meld_3
```{r,echo=FALSE,warning=FALSE,message=F}
tidycmprsk::crr(Surv(time_to_death_90days_cmp,time_to_death_status_90days_cmp) ~serious_ae+ age+race+hispanic+ admit_meld_3,failcode=1,cencode=0, data = final_master) %>% tbl_regression(exp = TRUE) %>% add_n()%>%
  bold_p(0.05) %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() 

```
#4. Multiviariate cmp:  time_to_death_status_90days_cmp failcode=1 death:serious_ae+ age+race+hispanic
```{r,echo=FALSE,warning=FALSE,message=F}
tidycmprsk::crr(Surv(time_to_death_90days_cmp,time_to_death_status_90days_cmp) ~serious_ae+ age+race+hispanic,failcode=1,cencode=0, data = final_master) %>% tbl_regression(exp = TRUE) %>% add_n()%>%
  bold_p(0.05) %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() 

```

# location_terli_day0 = 1 OR 2, logistic regression outcome: any_ae 
```{r}
final_master_location_terli_day0 <- final_master %>% filter(location_terli_day0 == 1| location_terli_day0 == 2)


final_master_location_terli_day0 %>%
  select(
   icu_admit, any_ae
  ) %>%
  is.na() %>%
  colSums()


final_master_location_terli_day0  %>%  select( icu_admit, any_ae
) %>%
  tbl_uvregression(
    method = glm,
    method.args = list(family = binomial),
    y =  icu_admit,
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  ) %>%
  bold_p(0.05) %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() %>% add_n()

```


# listed_transplant  == 1 Unviariate cmp:  time_to_death_status_90days_cmp failcode=2 transplant:   + missing number 
```{r,echo=,warning=FALSE}
final_master_listed_transplant  <- final_master %>% filter(listed_transplant  == 1)

final_master_listed_transplant %>%
  select(
    any_ae,
    serious_ae,
    hrs_responders_cat_2,
    day0_meld,
    meld_fda,
    time_to_death_90days_cmp,
    time_to_death_status_90days_cmp
  ) %>%
  is.na() %>%
  colSums()


final_master_listed_transplant   %>%  select( any_ae,  serious_ae,  hrs_responders_cat_2, day0_meld, meld_fda,
time_to_death_90days_cmp,time_to_death_status_90days_cmp) %>%
  tbl_uvregression(
    method = tidycmprsk::crr,
    method.args = list(failcode=2,cencode=0),
    y = Surv(time_to_death_90days_cmp,time_to_death_status_90days_cmp),
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  )
```