---
title: "Results"
author: "Tianqi Ouyang"
date: last-modified
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(tableone)
library(writexl)
library(gtsummary)
library(survival)
library(ggplot2)
library(readxl)
library(ggpubr)
library(splines)
library(xlsx)
library(lattice)
library(mice)
library(DescTools)
library(survminer)
library(finalfit)
library(tidycmprsk)
library(ggsurvfit)
library(car)
library(cobalt)
library(pammtools)
library(adjustedCurves)
library(rms)
library(riskRegression)
library(prodlim)
library(dplyr)
library(bodycompref)
library(musclerefdata)
library(transplantr)
library(sqldf)
library(table1)
```

```{r data-load, include=FALSE}
final_master <- read_xlsx("/Users/to909/Desktop/Terlipressin projects/Terlipressin/data/final_master_01282026.xlsx")

final_master <- final_master %>%
  mutate(across(everything(), ~{
    x_char <- as.character(.x)
    if (all(grepl("^[-+]?[0-9]*\\.?[0-9]+$", x_char[!is.na(x_char)]))) {
      as.numeric(x_char)
    } else {
      .x
    }
  }))

final_master <- final_master %>%
  mutate(
    BMI = aki_weight / (aki_height/100)^2,
    rrt_90days = ifelse(is.na(days_to_rrt), 0, ifelse(days_to_rrt <= 90, 1, 0)),
    lt_90days = ifelse(is.na(time_to_transplant), 0, ifelse(time_to_transplant <= 90, 1, 0)),
    rrt_status = case_when(
      (days_to_rrt - time_to_transplant) == 0 ~ 0,
      (days_to_rrt - time_to_transplant) < 0 ~ 1,
      (days_to_rrt - time_to_transplant) > 0 ~ 2
    ),
    abdominal_ae = if_else(terli_ae_abdpain == 1 | terli_ae_diarrhea == 1 | terli_ae_nausea == 1, 1, 0),
    serious_ae = if_else(terli_ae_respfail == 1 | terli_ae_brady == 1 | terli_7 == 1 | terli_8 == 1, 1, 0),
    meld_fda_value = coalesce(day0_meld, admit_meld, admit_meld_3),
    meld_fda = if_else(meld_fda_value > 35, 1, 0),
    aclf_grade = ifelse(aclf_grade == 0, 1, aclf_grade)
  )

# --- 04/07/2026 updates ---
# Coding fix: For chest_imaging=1, set missing effusion/edema to 0
final_master <- final_master %>%
  mutate(
    chest_imaging_effusion = if_else(chest_imaging == 1 & is.na(chest_imaging_effusion), 0, chest_imaging_effusion),
    chest_imaging_pulmedema = if_else(chest_imaging == 1 & is.na(chest_imaging_pulmedema), 0, chest_imaging_pulmedema)
  )

# New variables
final_master <- final_master %>%
  mutate(
    comb_effusion_edema = case_when(
      chest_imaging_effusion == 0 & chest_imaging_pulmedema == 0 ~ 0,
      chest_imaging_effusion == 1 & chest_imaging_pulmedema == 0 ~ 1,
      chest_imaging_effusion == 0 & chest_imaging_pulmedema == 1 ~ 2,
      chest_imaging_effusion == 1 & chest_imaging_pulmedema == 1 ~ 3
    ),
    ratio_albumin_b4terli = totalalbumin_before_terli / aki_weight,
    ratio_albumin_total = total_albumin / aki_weight
  )

final_master$time_to_death_status_90days_cmp <- as.factor(final_master$time_to_death_status_90days_cmp)
```

```{r table-one-function, include=FALSE}
# Custom render for continuous: Median [Q1, Q3]
render.continuous <- function(x, ...) {
  with(
    stats.apply.rounding(stats.default(x, ...), digits = 2),
    c("", `Median [Q1, Q3]` = sprintf("%s [%s, %s]", MEDIAN, Q1, Q3))
  )
}

# P-value function: Fisher's exact if any cell < 5, else chi-squared; Kruskal-Wallis for continuous
# Also tracks which test was used
pvalue <- function(x, ...) {
  y <- unlist(x)
  g <- factor(rep(1:length(x), times = sapply(x, length)))
  if (is.numeric(y)) {
    n_groups <- nlevels(droplevels(g))
    if (n_groups == 2) {
      p <- tryCatch(wilcox.test(y ~ g)$p.value, error = function(e) NA)
      test_used <<- c(test_used, "wilcox")
    } else {
      p <- tryCatch(kruskal.test(y ~ g)$p.value, error = function(e) NA)
      test_used <<- c(test_used, "kruskal")
    }
  }  else {
    tbl <- table(y, g)
    if (nrow(tbl) < 2 || ncol(tbl) < 2) {
      p <- NA
      test_used <<- c(test_used, "none")
    } else if (any(tbl < 5)) {
      p <- tryCatch(fisher.test(tbl)$p.value,
                    error = function(e) NA)
      test_used <<- c(test_used, "fisher")
    } else {
      p <- tryCatch(chisq.test(tbl)$p.value, error = function(e) NA)
      test_used <<- c(test_used, "chisq")
    }
  }
  pval <- if (is.na(p)) "" else sub("<", "&lt;", format.pval(p, digits = 3, eps = 0.001))
  c("", pval)
}

create_table_one <- function(all_vars, data, strata, output,
                             cat_vars = NULL, exact = NULL) {
  # Ensure strata is a factor
  data[[strata]] <- factor(data[[strata]])

  # Convert cat_vars to factor
  if (!is.null(cat_vars)) {
    for (v in intersect(cat_vars, names(data))) {
      data[[v]] <- factor(data[[v]])
    }
  }

  # Keep only vars that exist in data
  all_vars <- intersect(all_vars, names(data))

  # Build formula
  f <- as.formula(paste("~", paste(all_vars, collapse = " + "), "|", strata))

  # Reset test tracker
  test_used <<- character(0)

  # Render table1 with p-values and custom continuous render
  tbl <- table1(f, data = data, overall = "Overall",
                render.continuous = render.continuous,
                extra.col = list("P-value" = pvalue))

  # Also save CSV via tableone for backwards compatibility
  if (!is.null(cat_vars)) {
    numerical_vars <- setdiff(all_vars, cat_vars)
  } else {
    numerical_vars <- all_vars
  }
  T1 <- tableone::CreateTableOne(
    vars = all_vars, strata = strata, data = data,
    factorVars = cat_vars, addOverall = TRUE, includeNA = FALSE
  )
  T1_print <- print(T1, showAllLevels = TRUE, exact = exact,
                    nonnormal = numerical_vars, missing = TRUE,
                    quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
  write.csv(T1_print, file = output, row.names = TRUE)

  # Build footnote from tests used
  tests <- setdiff(unique(test_used), "none")
  labels <- c(
    wilcox  = "Mann-Whitney U test (continuous, 2 groups)",
    kruskal = "Kruskal-Wallis test (continuous, &ge;3 groups)",
    fisher  = "Fisher&rsquo;s exact test (categorical, any cell &lt; 5)",
    chisq   = "Chi-squared test (categorical)"
  )
  footnote <- paste(na.omit(labels[tests]), collapse = "; ")

  # Append footnote below the table
  htmltools::tagList(
    tbl,
    htmltools::tags$p(
      style = "font-size: 0.85em; color: #555; margin-top: 4px;",
      htmltools::HTML(paste0("Statistical tests: ", footnote))
    )
  )
}

# Initialize global test tracker
test_used <- character(0)
```

## 0. Missing-Data Check — 07/20/2026 Data Request {#missing-check}

Missingness for every variable involved in the 07/20/2026 request (cumulative dose comparison + multivariable AE models + Fine-Gray). Reported as `N missing / N total (% missing)`.

```{r missing-check, echo=TRUE}
miss_tbl <- function(df, vars, label) {
  vars <- intersect(vars, names(df))
  df %>%
    dplyr::select(dplyr::all_of(vars)) %>%
    dplyr::summarise(dplyr::across(dplyr::everything(),
                     ~ sprintf("%d / %d (%.1f%%)",
                               sum(is.na(.)), length(.),
                               100 * mean(is.na(.))))) %>%
    tidyr::pivot_longer(dplyr::everything(),
                        names_to = "variable", values_to = "missing") %>%
    dplyr::mutate(scope = label)
}

dose_vars <- paste0("total_dose_terli_day", 0:13)
covar_45  <- c("sex_male", "inr_admit", "scr_terli_day0", "chest_imaging_effusion")
covar_6   <- c("sex_male", "scr_terli_day0", "chest_imaging_effusion")
covar_7   <- c("any_ae", "age", "race", "hispanic", "admit_meld_3",
               "time_to_death_90days_cmp", "time_to_death_status_90days_cmp")

c1 <- final_master %>% dplyr::filter(terli_route_day0 == 1)
c2 <- final_master %>% dplyr::filter(terli_route_day0 == 2)

miss_out <- dplyr::bind_rows(
  miss_tbl(final_master, "terli_route_day0",                 "Full cohort: route flag"),
  miss_tbl(c1,           dose_vars,                          "Cohort route=1 (bolus): daily dose"),
  miss_tbl(c2,           dose_vars,                          "Cohort route=2 (infusion): daily dose"),
  miss_tbl(final_master, c("any_ae", covar_45),              "Model 4: any_ae + covariates"),
  miss_tbl(final_master, c("serious_ae", covar_45),          "Model 5: serious_ae + covariates"),
  miss_tbl(final_master, c("terli_ae_respfail", covar_6),    "Model 6: terli_ae_respfail + covariates"),
  miss_tbl(final_master, covar_7,                            "Model 7: Fine-Gray variables")
)

cat("Cohort sizes:\n",
    "  Full N          =", nrow(final_master), "\n",
    "  route=1 (bolus)     =", nrow(c1), "\n",
    "  route=2 (infusion)  =", nrow(c2), "\n\n")

knitr::kable(miss_out, caption = "Missingness for variables in the 07/20/2026 analysis request")
```

## 1. AE Incidence by Extended Criteria {#criteria-tables}

AE rates stratified by ACLF (`criteria_2`), sCr (`criteria_3`), and Hypoxemia (`criteria_4`) criteria.

```{r criteria-vars, include=FALSE}
ae_vars <- c("any_ae", "terli_ae_abdpain", "terli_ae_nausea", "terli_ae_diarrhea",
             "terli_ae_volover", "terli_ae_brady", "terli_7", "terli_8", "terli_ae_respfail",
             "serious_ae", "comb_effusion_edema", "ratio_albumin_b4terli", "ratio_albumin_total")

ae_cat_vars <- c("any_ae", "terli_ae_abdpain", "terli_ae_nausea", "terli_ae_diarrhea",
                 "terli_ae_volover", "terli_ae_brady", "terli_7", "terli_8", "terli_ae_respfail",
                 "serious_ae", "comb_effusion_edema")

exact_vars <- c("terli_ae_abdpain", "terli_ae_nausea", "terli_ae_diarrhea",
                "terli_ae_volover", "terli_ae_brady", "terli_7", "terli_8")
```

### ACLF (criteria_2)
```{r criteria-2}
create_table_one(
  all_vars = ae_vars, cat_vars = ae_cat_vars,
  data = final_master, exact = exact_vars,
  strata = "criteria_2",
  output = '/Users/to909/Desktop/Terlipressin projects/Adverse event/results/table1_criteria_2.csv'
)
```

### sCr (criteria_3)
```{r criteria-3}
create_table_one(
  all_vars = ae_vars, cat_vars = ae_cat_vars,
  data = final_master, exact = exact_vars,
  strata = "criteria_3",
  output = '/Users/to909/Desktop/Terlipressin projects/Adverse event/results/table1_criteria_3.csv'
)
```

### Hypoxemia (criteria_4)
```{r criteria-4}
create_table_one(
  all_vars = ae_vars, cat_vars = ae_cat_vars,
  data = final_master, exact = exact_vars,
  strata = "criteria_4",
  output = '/Users/to909/Desktop/Terlipressin projects/Adverse event/results/table1_criteria_4.csv'
)
```

## 2. Baseline Characteristics by AE Type {#baseline-tables}

```{r baseline-vars, include=FALSE}
categorical_vars <- c(
  "admit_admitservice", "admit_admitroute", "sex_male", "race", "hispanic",
  "insurance_type", "cirrhosis_etiology", "etoh_hepatitis", "hegrade_admit",
  "admit_ascites", "htn_medical_hx", "dm_medical_hx", "ckd_medical_hx",
  "cad_medical_hx", "nsbb_preadmission", "nsbb_type_preadmission",
  "ace_arb_preadmission", "nsaid_preadmission", "diuretics_preadmission",
  "sbp_ppx_preadmission", "lactulose_preadmission", "rifaximin_preadmission",
  "mido_preadmission", "ascites_previous_comp", "refascites_previous_comp",
  "he_previous_comp", "sbp_previous_comp", "nonsbp_previous_comp",
  "phtn_bleed_previous_comp", "hcc_previous_comp", "tips_previous_comp",
  "baseline_scr_outpt", "aki_previous_90days", "transfer_aki",
  "aki_precipitant_nsbpinf", "he_aki", "rrt_aki", "mv_aki", "pressors_aki",
  "chest_imaging", "chest_imaging_pulmedema", "chest_imaging_effusion",
  "ecg", "echo", "albumin_before_terli", "crystalloid_before_terli",
  "nsbb_before_terli", "mido_before_terli", "oct_before_terli",
  "ne_before_terli", "gihep_consult", "neph_consult", "pall_consult",
  "code_status", "ua_hrs", "ua_na_hrs", "ua_cr_hrs", "ua_urea_hrs",
  "ua_granularcasts_hrs", "ua_epi_cells_hrs", "location_terli_day0",
  "terli_route_day0", "tele_terli_day0", "he_terli_day0",
  "infection_terli_day0", "typeinfection_terli_day0", "hrs_responders",
  "dosechange_terli_day1", "dosechange_terli_day2", "dosechange_terli_day3",
  "dosechange_terli_day4", "dosechange_terli_day5", "dosechange_terli_day6",
  "dosechange_terli_day7", "dosechange_terli_day8", "dosechange_terli_day9",
  "dosechange_terli_day10", "dosechange_terli_day11", "dosechange_terli_day12",
  "discont_aki_stage", "aclf_grade", "HRS_Criteria", "criteria_1",
  "criteria_2", "criteria_3", "criteria_4", "akin", "transplant_slkt",
  "transplant_lt", "rrt", "rrt_2", "rrt_modality", "rrt_continuous",
  "hd_discharge", "disposition", "liver_f", "brain_f", "coagulation_f",
  "circulatory_f", "respiratory_f", "kidney_f", "baseline_scr_120to365",
  "aki_events_previous90days", "prbc_trans_before_terli",
  "time_to_death_status", "terli_ae_abdpain", "terli_ae_nausea",
  "terli_ae_diarrhea", "terli_ae_volover", "terli_ae_brady", "terli_7",
  "terli_8", "terli_ae_respfail", "any_ae", "listed_transplant",
  "dosechange_status", "hrs_rebound", "race_2", "HRS_reversal_new",
  "hrs_responders_cat_2", "time_to_death_status_90days",
  "time_to_death_status_90days_cmp", "ckd_baseline", "HRS_criteria_3"
)

numerical_vars <- c(
  "age", "scr_admit", "tbili_admit", "inr_admit", "sodium_admit",
  "albumin_admit", "lactate_admit", "nsbb_dose_preadmission",
  "mido_dose_preadmission", "baseline_scr", "baseline_scr_days_to_admit",
  "transfer_aki_days_toadmit", "days_admit_to_aki", "aki_weight",
  "aki_height", "aki_hr", "aki_sbp", "aki_dbp", "aki_map", "aki_pao2",
  "fio2_aki", "scr_aki", "bun_aki", "albumin_aki", "inr_aki", "wbc_aki",
  "plt_aki", "ecg_qtc", "days_to_echo", "echo_ef", "echo_lateral_e",
  "echo_trpeak", "echo_gls", "echo_earatio", "echo_rvsp", "echo_pasp",
  "echo_lavi", "echo_tr", "totalalbumin_before_terli",
  "albvolume_before_terli", "saline_vol_before_terli",
  "lr_volume_before_terli", "other_cryst_before_terli",
  "prbc_qaunt_before_terli", "totalvolume_before_terli",
  "total_uo_before_terli", "days_first_mido_before_terli",
  "days_total_mido_before_terli", "mido_maxdose_before_terli",
  "days_first_oct_before_terli", "days_total_oct_before_terli",
  "oct_maxdose_before_terli", "days_first_ne_before_terli",
  "days_total_ne_before_terli", "days_to_gihep_consult",
  "days_to_neph_consult", "days_to_pall_consult", "days_ua_hrs",
  "days_ua_na_hrs", "ua_na_value_hrs", "days_ua_cr_hrs", "ua_cr_value_hrs",
  "days_ua_urea_hrs", "ua_urea_value_hrs", "feurea_hrs", "ua_rbc_hrs",
  "uaprotein_hrs", "temp_terli_day0", "hr_terli_day0", "pao2_terli_day0",
  "fio2_terli_day0", "sbp_terli_day0_time0", "map_terli_day0_time0",
  "chloride_terli_day0", "bicarb_terli_day0", "scr_terli_day0",
  "bun_terli_day0", "ast_terli_day0", "alt_terli_day0", "bili_terli_day0",
  "alb_terli_day0", "inr_terli_day0", "wbc_terli_day0", "hgb_terli_day0",
  "Discont_Day", "na_terli_day0", "days_to_terli_initiation_day0",
  "clif_c_score", "admit_meld", "admit_meld_na", "admit_meld_3",
  "admit_meld_3_raw", "day0_meld", "day0_meld_na", "day0_meld_3",
  "time_to_transplant", "time_to_death_status", "time_to_death_90days",
  "cystatinc_admit", "cystatinc_aki", "cystatinc_terli_day0",
  "urine_pc", "fena", "fena_hrs", "total_albumin"
)

categorical_vars <- intersect(categorical_vars, names(final_master))
numerical_vars <- intersect(numerical_vars, names(final_master))
all_vars <- c(numerical_vars, categorical_vars)

final_master <- final_master %>% mutate_at(categorical_vars, as.factor)
```

### Any AE
```{r table1-any-ae}
create_table_one(
  all_vars = all_vars, cat_vars = categorical_vars,
  data = final_master, strata = "any_ae",
  output = '/Users/to909/Desktop/Terlipressin projects/Adverse event/results/table1_any_ae.csv'
)
```

### Abdominal AE
```{r table1-abdominal-ae}
create_table_one(
  all_vars = all_vars, cat_vars = categorical_vars,
  data = final_master, strata = "abdominal_ae",
  output = '/Users/to909/Desktop/Terlipressin projects/Adverse event/results/table1_abdominal_ae.csv'
)
```

### Serious AE
```{r table1-serious-ae}
create_table_one(
  all_vars = all_vars, cat_vars = categorical_vars,
  data = final_master, strata = "serious_ae",
  output = '/Users/to909/Desktop/Terlipressin projects/Adverse event/results/table1_serious_ae.csv'
)
```

### Respiratory Failure
```{r table1-respfail}
create_table_one(
  all_vars = all_vars, cat_vars = categorical_vars,
  data = final_master, strata = "terli_ae_respfail",
  output = '/Users/to909/Desktop/Terlipressin projects/Adverse event/results/table1_terli_ae_respfail.csv'
)
```

## 2b. AE Cross-Tabulations (04/07/2026 Update) {#ae-cross-tables}

### SAE (Serious AE as column)
```{r table-sae}
sae_vars <- c("any_ae", "terli_ae_abdpain", "terli_ae_nausea", "terli_ae_diarrhea",
              "terli_ae_volover", "terli_ae_brady", "terli_7", "terli_8",
              "terli_ae_respfail", "comb_effusion_edema", "ratio_albumin_b4terli", "ratio_albumin_total")

sae_cat_vars <- c("any_ae", "terli_ae_abdpain", "terli_ae_nausea", "terli_ae_diarrhea",
                  "terli_ae_volover", "terli_ae_brady", "terli_7", "terli_8",
                  "terli_ae_respfail", "comb_effusion_edema")

create_table_one(
  all_vars = sae_vars, cat_vars = sae_cat_vars,
  data = final_master, strata = "serious_ae",
  output = '/Users/to909/Desktop/Terlipressin projects/Adverse event/results/table1_serious_ae_cross.csv'
)
```

### Respiratory Failure (with new variables)
```{r table-respfail-updated}
respfail_vars <- c("any_ae", "terli_ae_abdpain", "terli_ae_nausea", "terli_ae_diarrhea",
                   "terli_ae_volover", "terli_ae_brady", "terli_7", "terli_8",
                   "serious_ae", "comb_effusion_edema", "ratio_albumin_b4terli", "ratio_albumin_total")

respfail_cat_vars <- c("any_ae", "terli_ae_abdpain", "terli_ae_nausea", "terli_ae_diarrhea",
                       "terli_ae_volover", "terli_ae_brady", "terli_7", "terli_8",
                       "serious_ae", "comb_effusion_edema")

create_table_one(
  all_vars = respfail_vars, cat_vars = respfail_cat_vars,
  data = final_master, strata = "terli_ae_respfail",
  output = '/Users/to909/Desktop/Terlipressin projects/Adverse event/results/table1_terli_ae_respfail_updated.csv'
)
```

## 3. ACLF Grade by AE Status {#aclf-table}

```{r aclf-by-ae}
table1(~aclf_grade | any_ae, final_master)
```

## 4. Logistic Regression — Predictors of Any AE {#logistic}

### Univariable
```{r logistic-uv}
final_master <- final_master %>% mutate_at(c("aclf_grade", "respiratory_f"), as.factor)

final_master %>%
  select(any_ae, scr_terli_day0, respiratory_f, aclf_grade) %>%
  tbl_uvregression(
    method = glm,
    method.args = list(family = binomial),
    y = any_ae,
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  ) %>%
  modify_caption("**Univariable logistic regression for any adverse event**") %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  add_n()
```

### Multivariable (04/07/2026 Update): sex_male + inr_admit + akin + chest_imaging_effusion
```{r logistic-mv-updated}
final_master <- final_master %>% mutate_at(c("akin"), as.factor)

glm(any_ae ~ sex_male + inr_admit + akin + chest_imaging_effusion,
    family = binomial, data = final_master) %>%
  tbl_regression(exponentiate = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  modify_caption("**Multivariable logistic regression for any adverse event (updated)**")
```

## 5. Competing Risk Models — 90-Day Mortality {#cmprisk}

### Univariable (Fine-Gray, failcode=1: death)
```{r cmprisk-uv}
final_master %>%
  select(any_ae, terli_ae_abdpain, terli_ae_nausea, terli_ae_diarrhea,
         terli_ae_volover, terli_ae_brady, terli_7, terli_8,
         terli_ae_respfail, serious_ae,
         time_to_death_90days_cmp, time_to_death_status_90days_cmp) %>%
  tbl_uvregression(
    method = tidycmprsk::crr,
    method.args = list(failcode = 1, cencode = 0),
    y = Surv(time_to_death_90days_cmp, time_to_death_status_90days_cmp),
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  )
```

### Model 1: any_ae + age + race + hispanic + admit_meld_3 (updated 04/07/2026)
```{r cmprisk-mv1}
tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, time_to_death_status_90days_cmp) ~
    any_ae + age + race + hispanic + admit_meld_3,
  failcode = 1, cencode = 0, data = final_master
) %>%
  tbl_regression(exp = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

### Model 2: any_ae + age + race + hispanic
```{r cmprisk-mv2}
tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, time_to_death_status_90days_cmp) ~
    any_ae + age + race + hispanic,
  failcode = 1, cencode = 0, data = final_master
) %>%
  tbl_regression(exp = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

### Model 3: serious_ae + age + race + hispanic + admit_meld_3
```{r cmprisk-mv3}
tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, time_to_death_status_90days_cmp) ~
    serious_ae + age + race + hispanic + admit_meld_3,
  failcode = 1, cencode = 0, data = final_master
) %>%
  tbl_regression(exp = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

### Model 4: serious_ae + age + race + hispanic
```{r cmprisk-mv4}
tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, time_to_death_status_90days_cmp) ~
    serious_ae + age + race + hispanic,
  failcode = 1, cencode = 0, data = final_master
) %>%
  tbl_regression(exp = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

## 6. Subgroup: ICU Admission {#icu}

Cohort: patients with `location_terli_day0 == 1` or `2` (floor or step-down at terlipressin initiation).

```{r icu-subgroup}
final_master_location <- final_master %>%
  filter(location_terli_day0 == 1 | location_terli_day0 == 2)

final_master_location %>%
  select(icu_admit, any_ae) %>%
  tbl_uvregression(
    method = glm,
    method.args = list(family = binomial),
    y = icu_admit,
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  ) %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  add_n()
```

## 7. Subgroup: Transplant-Listed Patients {#transplant}

Cohort: `listed_transplant == 1`. Competing risk model with failcode=2 (transplant event).

```{r transplant-subgroup}
final_master_listed <- final_master %>% filter(listed_transplant == 1)

final_master_listed %>%
  select(any_ae, serious_ae, hrs_responders_cat_2, day0_meld, aclf_grade,
         meld_fda, time_to_death_90days_cmp, time_to_death_status_90days_cmp) %>%
  tbl_uvregression(
    method = tidycmprsk::crr,
    method.args = list(failcode = 2, cencode = 0),
    y = Surv(time_to_death_90days_cmp, time_to_death_status_90days_cmp),
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  )
```

## 8. AE Data Request (07/20/2026) — Complete-Case Analyses {#request-0720}

Complete-case analyses (no imputation for now — to be updated later). **Models 4–6 are fit on the full analytic cohort**; the *n* in the request (42 / 32 / 23) is the number of **events** (outcome = 1), not a restricted cohort — restricting to event-positive patients would make the outcome constant and the model unfit.

::: {.callout-warning}
**Complete-case shrinks EPV below the request's estimate.** Dropping rows with any missing covariate reduces the effective events, so events-per-variable falls versus the planned 10.5 / 8.0 / 7.7:

| Model | Outcome | Full-cohort events | Complete-case N (events) | Effective EPV |
|:--|:--|:--:|:--:|:--:|
| 4 | `any_ae` | 42 | 191 (32) | **8.0** |
| 5 | `serious_ae` | 32 | 191 (23) | **5.8** |
| 6 | `terli_ae_respfail` | 23 | 196 (17) | **5.7** |

Models 5 and 6 (EPV < 6) are prone to overfitting and unstable estimates — interpret ORs cautiously. Multiple imputation (planned next) would recover the dropped events and restore EPV. Each model caption below recomputes N / events / EPV at render time; a separate diagnostics table flags any model showing quasi-complete separation.
:::

### 8.1 Cumulative terlipressin dose by day-0 route {#dose-sum}

Per-patient cumulative dose = sum of the daily total terlipressin dose, day 0 through day 13, missing daily dose treated as 0 (`na.rm = TRUE`). Cohorts: `terli_route_day0 == 1` (Task 1, IV bolus) and `== 2` (Task 2, continuous infusion).

::: {.callout-important}
**Column-name correction.** The request listed `total_dose_terli_day0 … total_dose_terli_day13`, but the master dataset only has `total_dose_terli_day0` under that name — the day 1–13 daily totals are stored as **`total_terli_day1 … total_terli_day13`** (the word "dose" is dropped for days 1–13). This section uses the correct columns. Summing only `total_dose_terli_day*` (as literally written) would have collapsed the "cumulative" dose to just the day-0 value.
:::

```{r dose-cumulative}
#| code-fold: true
#| code-summary: "Show code"
# Daily total terlipressin dose: day 0 uses total_dose_terli_day0;
# days 1-13 use total_terli_day{N} (naming differs in the export).
dose_vars_all <- c("total_dose_terli_day0", paste0("total_terli_day", 1:13))
dose_missing  <- setdiff(dose_vars_all, names(final_master))
dose_vars     <- intersect(dose_vars_all, names(final_master))
if (length(dose_missing) > 0) message("Dose columns not found: ", paste(dose_missing, collapse = ", "))

# route may be a factor by this point — coerce back to numeric safely
route_num <- suppressWarnings(as.numeric(as.character(final_master$terli_route_day0)))

dose_df <- final_master %>%
  mutate(
    route_num = route_num,
    cumulative_dose = rowSums(
      across(all_of(dose_vars), ~ suppressWarnings(as.numeric(as.character(.)))),
      na.rm = TRUE
    )
  ) %>%
  filter(route_num %in% c(1, 2)) %>%
  mutate(route = factor(route_num, levels = c(1, 2),
                        labels = c("IV bolus (route=1)", "Continuous infusion (route=2)")))

dose_summary <- dose_df %>%
  group_by(route) %>%
  summarise(
    N              = n(),
    `Mean +/- SD`  = sprintf("%.1f +/- %.1f", mean(cumulative_dose), sd(cumulative_dose)),
    `Median [IQR]` = sprintf("%.1f [%.1f, %.1f]", median(cumulative_dose),
                             quantile(cumulative_dose, .25), quantile(cumulative_dose, .75)),
    `Min, Max`     = sprintf("%.1f, %.1f", min(cumulative_dose), max(cumulative_dose)),
    `Group total`  = sprintf("%.1f", sum(cumulative_dose)),
    .groups = "drop"
  )

knitr::kable(dose_summary,
             caption = "Per-patient cumulative terlipressin dose (day 0-13) by day-0 route")

# --- Data-quality flag: implausible daily doses (typical terli <= ~12 mg/day) ---
flag <- dose_df %>%
  mutate(across(all_of(dose_vars), ~ suppressWarnings(as.numeric(as.character(.))))) %>%
  rowwise() %>%
  mutate(max_daily = max(c_across(all_of(dose_vars)), na.rm = TRUE)) %>%
  ungroup() %>%
  filter(max_daily > 50) %>%
  select(route, any_of(c("record_id", "study_id")), max_daily)
if (nrow(flag) > 0) {
  cat("DATA-QUALITY FLAG - daily dose > 50 mg (likely entry error; inflates the mean/t-test):\n")
  print(as.data.frame(flag))
}
```

### 8.2 Comparison of cumulative dose between routes (Task 3) {#dose-compare}

Normality screened per group (Shapiro–Wilk); both parametric (Welch t) and non-parametric (Mann–Whitney U) tests reported.

- **Both groups are right-skewed and non-normal → Mann–Whitney U is the primary test.** The Welch t-test is unreliable here because two extreme values (day-0 = 701.7, day-2 = 425; see the data-quality flag above) inflate the bolus-group mean and hide the difference.
- **Duration confounding:** cumulative dose reflects total exposure, not dosing intensity — a longer infusion accrues more drug at equal intensity. A dose-per-treatment-day comparison can be added later.

```{r dose-compare}
#| code-fold: true
#| code-summary: "Show code"
sw <- dose_df %>%
  group_by(route) %>%
  summarise(
    n = n(),
    shapiro_p = tryCatch(shapiro.test(cumulative_dose)$p.value,
                         error = function(e) NA_real_),
    .groups = "drop"
  )

vt <- tryCatch(var.test(cumulative_dose ~ route, data = dose_df), error = function(e) NULL)
tt <- tryCatch(t.test(cumulative_dose ~ route, data = dose_df), error = function(e) NULL)   # Welch
wt <- tryCatch(wilcox.test(cumulative_dose ~ route, data = dose_df), error = function(e) NULL)

cat("Shapiro-Wilk normality by group:\n"); print(sw)
cat(sprintf("\nEqual-variance F-test:  p = %s\n",
            ifelse(is.null(vt), "NA", format.pval(vt$p.value, digits = 3))))
cat(sprintf("Welch t-test:           p = %s  (mean difference = %.1f)\n",
            ifelse(is.null(tt), "NA", format.pval(tt$p.value, digits = 3)),
            ifelse(is.null(tt), NA_real_, unname(tt$estimate[1] - tt$estimate[2]))))
cat(sprintf("Mann-Whitney U test:    p = %s\n",
            ifelse(is.null(wt), "NA", format.pval(wt$p.value, digits = 3))))

primary <- if (all(sw$shapiro_p > 0.05, na.rm = TRUE))
  "Welch t-test (both groups pass normality)" else
  "Mann-Whitney U (>=1 group non-normal) - recommended primary given small N"
cat(sprintf("\nSuggested primary test: %s\n", primary))
```

### 8.3 Model 4 — Multivariable logistic: Any AE {#model4}

`any_ae ~ sex_male + inr_admit + scr_terli_day0 + chest_imaging_effusion` (complete-case).

```{r model4-anyae}
#| code-fold: true
#| code-summary: "Show code"
m4 <- glm(any_ae ~ sex_male + inr_admit + scr_terli_day0 + chest_imaging_effusion,
          family = binomial, data = final_master)

tbl_regression(m4, exponentiate = TRUE,
               pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  modify_caption(sprintf(
    "**Model 4 - Any AE (complete-case): N = %d, events = %d, EPV = %.1f**",
    stats::nobs(m4), sum(m4$y), sum(m4$y) / 4))
```

### 8.4 Model 5 — Multivariable logistic: Serious AE {#model5}

`serious_ae ~ sex_male + inr_admit + scr_terli_day0 + chest_imaging_effusion` (complete-case).

```{r model5-seriousae}
#| code-fold: true
#| code-summary: "Show code"
m5 <- glm(serious_ae ~ sex_male + inr_admit + scr_terli_day0 + chest_imaging_effusion,
          family = binomial, data = final_master)

tbl_regression(m5, exponentiate = TRUE,
               pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  modify_caption(sprintf(
    "**Model 5 - Serious AE (complete-case): N = %d, events = %d, EPV = %.1f**",
    stats::nobs(m5), sum(m5$y), sum(m5$y) / 4))
```

### 8.5 Model 6 — Multivariable logistic: Respiratory AE / Volume Overload {#model6}

`terli_ae_respfail ~ sex_male + scr_terli_day0 + chest_imaging_effusion` (3 covariates, complete-case).

```{r model6-respfail}
#| code-fold: true
#| code-summary: "Show code"
m6 <- glm(terli_ae_respfail ~ sex_male + scr_terli_day0 + chest_imaging_effusion,
          family = binomial, data = final_master)

tbl_regression(m6, exponentiate = TRUE,
               pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  modify_caption(sprintf(
    "**Model 6 - Respiratory AE (complete-case): N = %d, events = %d, EPV = %.1f**",
    stats::nobs(m6), sum(m6$y), sum(m6$y) / 3))
```

### 8.6 Model diagnostics (complete-case) {#model-diagnostics}

Effective sample, events, EPV, and a crude separation flag (any |β| > 10 or SE > 10, which signals quasi-complete separation — likely with few events + binary predictors). If flagged, treat ORs/CIs as unstable and consider Firth penalized logistic (`logistf`) as a later sensitivity analysis.

```{r model-diagnostics}
#| code-fold: true
#| code-summary: "Show code"
diag_row <- function(m, k, label) {
  b <- coef(summary(m))
  data.frame(
    Model                 = label,
    N                     = stats::nobs(m),
    Events                = sum(m$y),
    Covariates            = k,
    EPV                   = round(sum(m$y) / k, 1),
    `Separation?`         = ifelse(any(abs(b[, "Estimate"]) > 10 | b[, "Std. Error"] > 10,
                                        na.rm = TRUE), "YES - check", "no"),
    check.names = FALSE
  )
}
knitr::kable(
  rbind(diag_row(m4, 4, "4 - any_ae"),
        diag_row(m5, 4, "5 - serious_ae"),
        diag_row(m6, 3, "6 - terli_ae_respfail")),
  row.names = FALSE,
  caption = "Complete-case model diagnostics")
```

### 8.7 Model 7 — Fine–Gray competing-risk model {#model7}

`any_ae + age + race + hispanic + admit_meld_3`, failcode = 1 (death), transplant competing. **Identical specification to §5 Model 1** — re-presented here so the 07/20 request is self-contained.

```{r model7-finegray}
#| code-fold: true
#| code-summary: "Show code"
tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, time_to_death_status_90days_cmp) ~
    any_ae + age + race + hispanic + admit_meld_3,
  failcode = 1, cencode = 0, data = final_master
) %>%
  tbl_regression(exp = TRUE, pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  modify_caption("**Model 7 - Fine-Gray, 90-day mortality (failcode = 1, death); identical to Sec. 5 Model 1**")
```
